MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-25 chapters/MetaTracts.tex
figures done
Local text commit

The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On one hand, ultrasonic testing provides a quick and cost-efficient overview of the material. On the other hand, it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component. Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen. In the typical scanning geometry of industrial XCT, the specimen is placed on a rotary table between X-ray source and detector. X-rays are emitted by the source in the form of a conic beam. When passing through the specimen the X-rays are attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below 500 $nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

@@

\subsection {Fiber Tracking in Diffusion Tensor Imaging}
\label{subsec:fiberEx} 
A basic assumption in DTI analysis is that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel~\cite{Basser2002, Basser2000, Mori1999, Mori2002}. Continuous tracts are created by propagating virtual particles along fiber directions until they reach some termination criterion. 
This is usually done by solving a second-order Runge-Kutta integration. The principal diffusion direction at each discrete location is interpolated to form a continuous velocity field. 
Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers. Basser et al.~\cite{Basser2002,Basser2000} proposed that white matter tracts could be represented as 3D curves in space. They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data. 
Mori et al.~\cite{Mori1999,Mori2002} divided reconstruction techniques into line-propagation or energy minimization techniques. In line propagation approaches, trajectories are computed based on local neighborhoods and in energy minimization approaches the most  favorable trajectory connecting two given endpoints is selected.

@@

\begin{figure}[tb]
\centering
	\includegraphics[width=\linewidth]{images_pvis/figure5.png}
	%\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
	\caption{(a) MetaTracts color-coded according to their mean orientations. (b) length distribution of individual MetaTracts for a particular bundle (unit for length is grid cube edge length:  $2\mu m$). A single fiber bundle is made of MetaTracts of varying length}
	\label{fig:length_distribution}
\end{figure}

@@

\subsection{Choice of clustering techniques}
As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
%\begin{figure}[tb]
%\centering
%		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
%		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
%	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
%	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
%	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
%	\label{fig:crop-16-decomp}
%\end{figure}  
%  
Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering. As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers. When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
%
On the other hand, Figure.~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step for 10, 15, and 20 clusters. In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
\begin{figure}[tb]
	\centering
	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
	\caption{Applying only hierarchical (a) and K-means (b) clustering methods of MetaTracts for various numbers of clusters.}
	\label{fig:comparison}
\end{figure}

@@

\begin{figure}
\centering
		\includegraphics[width=\linewidth]{images_pvis/figure7}
		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(a) Voxelization of data set 1; (b) a single slice of the data set; (c) two of the extracted meshes together; (d) the meshes rendered separately.}
	\label{fig:crop-16-decomp}
\end{figure}

>>>>

The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On one hand, ultrasonic testing provides a quick and cost-efficient overview of the material. On the other hand, it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component. Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen. In the typical scanning geometry of industrial XCT, the specimen is placed on a rotary table between X-ray source and detector. When passing through the specimen the X-rays are attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below 500 $nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

@@

\subsection {Fiber Tracking in Diffusion Tensor Imaging}
\label{subsec:fiberEx} 
A basic assumption in DTI analysis is that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel~\cite{Basser2002, Basser2000, Mori1999, Mori2002}. Continuous tracts are created by propagating virtual particles along fiber directions until they reach some termination criterion. 
This is usually done by solving a Runge-Kutta integration (typically second or fourth order). The principal diffusion direction at each discrete location is interpolated to form a continuous velocity field. 
Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers. Basser et al.~\cite{Basser2002,Basser2000} proposed that white matter tracts could be represented as 3D curves in space. They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data. 
Mori et al.~\cite{Mori1999,Mori2002} divided reconstruction techniques into line-propagation or energy minimization techniques. In line propagation approaches, trajectories are computed based on local neighborhoods and in energy minimization approaches the most  favorable trajectory connecting two given endpoints is selected.

@@

\begin{figure}[tb]
\centering
	\includegraphics[width=\linewidth]{images_pvis/figure5.pdf}
	%\includegraphics[width=0.45\textwidth,  trim = 0mm 100mm 0mm 50mm, clip]{imagesMT2014/image_length}
	\caption{(a) MetaTracts color-coded according to their mean orientations. (b) length distribution of individual MetaTracts for a particular bundle (unit for length is grid cube edge length:  $2\mu m$). A single fiber bundle is made of MetaTracts of varying length}
	\label{fig:length_distribution}
\end{figure}

@@

\subsection{Choice of clustering techniques}
As opposed to the two phase clustering we experimented with using proximity alone and using single step clustering of MetaTracts.
%\begin{figure}[tb]
%\centering
%		\includegraphics[width=0.4\textwidth]{images_pvis/figure7}
%		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
%	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
%	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
%	\caption{(A) Volume rendering of data set Fig.~\ref{fig:data-char}A.(B) shows a single slice of the data set. (C) shows two of the extracted meshes together. (D) shows the meshes separately.}
%	\label{fig:crop-16-decomp}
%\end{figure}  
%  
Figure~\ref{fig:comparison}a shows the MetaTracts that are hierarchically clustered directly using proximity alone into 10, 15, and 20 clusters (ground truth is 11 clusters) without first performing orientation clustering. As discussed above, single linkage hierarchical clustering in the presence of overlapping fibers tends to create large clusters and small, low-cardinality outliers. When number of clusters ($h$) is 10 two large incorrect clusters are generated and the rest are outliers. As $h$ increases, some appropriate bundles start to form. Even at $h=20$, the top bundles all clusters together into one large incorrectly identified bundle.
%
On the other hand, Figure.~\ref{fig:comparison}b shows the MetaTracts, clustered using K-means clustering after being embedded in a $m$-dimensional (lower) space using proximity alone, without a second hierarchical clustering step for 10, 15, and 20 clusters. In the lower dimensional space the spatial context is lost and fibers which are in reality far away are grouped together.
\begin{figure}[tb]
	\centering
	\includegraphics[width=\linewidth]{images_pvis/comparison_all.pdf}
	\caption{Applying only hierarchical (a) and K-means (b) clustering methods to MetaTracts for various numbers of clusters.}
	\label{fig:comparison}
\end{figure}

@@

\begin{figure}
\centering
		\includegraphics[width=\linewidth]{images_pvis/figure8.pdf}
		%\includegraphics[width=0.5\textwidth, trim = 0mm 70mm 0mm 20mm, clip]{imagesMT2014/image7b.pdf}
	%\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	%\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(a) Voxelization of data set 1; (b) a single slice of the data set; (c) two of the extracted meshes together; (d) the meshes rendered separately.}
	\label{fig:crop-16-decomp}
\end{figure}
